\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage[linguistics]{forest}
\usepackage{listings}

\title{Chapter 20 CLRS: Graphs}
\author{Rafael Juarez Laureano}

\begin{document}
\maketitle

\section{Representation of graphs}

\subsection{20.1-1}
Given an adjacency-list representation of a direct graph, how long does it take to compute the out-degree of every vertex?
How long does it take to compute the in-degrees?
Answer:
\begin{itemize}
    \item 
        We have to go through the adjacency list looking for a node, and taking this row, this give us a complexity:
        
        $\mathcal{O}(|V| + |E|)$
    
    \item 
        We have to count every appearance of the selected node throughout the adjacency-list:
        
        $\mathcal{O}(|V| + |E|)$
\end{itemize}
 
\subsection{20.1-2}
Given an ajacency-list representacion for a complete binary tree on 7 vertices.Give an equivalent adjacency-matrix representation. Assume that the edges are undirected and that vertices are numbered from 1 to 7 as in a binary heap.

Tree:
\begin{forest}
[1 
    [2 [[4] [5]]] 
    [3 [[6] [7]]]
]
\end{forest}
\begin{itemize}
    \item 
        Adjacency-Matriz 
        $\begin{bmatrix}
        NIL & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
        1 & 0 & 1 & 1 & 0 & 0 & 0 & 0 \\
        2 & 1 & 0 & 0 & 1 & 1 & 0 & 0 \\
        3 & 1 & 0 & 0 & 0 & 0 & 1 & 1 \\
        4 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
        5 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
        6 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
        7 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
        \end{bmatrix}$
    \item 
        Adjacency-List
        $\begin{bmatrix}
        1: & 2 & 3 \\
        2: & 1 & 4 & 5 \\
        3: & 1 & 6 & 7 \\
        4: & 2 \\
        5: & 2  \\
        6: & 3 \\
        7: & 3 \\
        \end{bmatrix}$
\end{itemize}

\subsection{20.1-3}
The transpose of a directed graph G=(V,E) is the graph 
$G^T = (V,E^T)$, where $E^T = {(u, v) \in VxV : (u,v) \in E}$
That is, $G^T$ is G with all its edges reversed. Describe efficient algorithms for computing $G^T$ from G, for both the adjacency-list and adjacency-matrix representation of G. Analyze the running times of your algorithms. 
\\
Adjacency-list:
\begin{verbatim}
let adjT[1..v] the new empty transpose adjacency-list 
for each v vertex in adj:
    for each u vertex in adf[v]:
        adjT[u].insert(v)
\end{verbatim}
Time Complexity: $O(|E| + |V|)$
\\ 
Auxiliary Space: $O(|E| + |E|)$
\\
Adjacenct-matrix:
Going through the main diagonal and swap every element below the superior diagonal.
\begin{verbatim}
for each vertex v in adj:
    for each u column in adj[u] | u >= v :
            swap(adj[v][u], adj[u][v])
\end{verbatim}
Time Complexity: $O(|V|^2)$
\\ 
Auxiliary Space: $O(1)$
\\
\subsection{20.1-4}
Given an adjacency-list representation of a multigraph G=(V,E), describe an O(V+E)-time algorithm to compute the adjacency-list representation of the "equivalent" undirected graph G'=(V,E'), where E' consists of the edges in E with all multiple edges between two vertices replaced by a single edge and with all self-loops removed.

\begin{verbatim}
let adj'[1..V] a new adjacency-list
for v vertex in adj:
    let visited be a |V| 0-initialized array
    for u vertex in adj[v]:
        if u != v and visited[u] != 1:
            visited[u] = 1
            adj'[v].insert(u)
\end{verbatim}

\subsection{20.1-4}
The square of a directed) G=(V,E) is the graph ) $G^2=(V,E^2)$ such that  $(u,v) \in E^2$ if and only if G contains a path with at most two edges between u and v. Describe efficient algorithms for computing $G^2$ from G for both the adjacency-list and adjacency-matrix representations of G. Analyze the running times of your algorithms.
\\
For Adjacency-List
\begin{verbatim}
let adj'[1..v] a new Adjacency-list
for v vertex in adj:
    for u vertex in adj[v]:
        adj'[v].insert(u)
        for w vertex in adj[u]:
            adj'[v].insert(w)
\end{verbatim}
\begin{itemize}
\item Time Complexity: $O(|V|*|E|)$
\item Auxiliary Space: $O(|V| + |E|)$
\end{itemize}
For adjacency-Matrix
\begin{verbatim}
let adj'[1..v][1..v] a new Adjacency-matrix
for v row in adj:
    for u column in adj[v]:
        adj'[v, u] = 1
        for w vertex in adj[u]:
            adj'[v, w] = 1
\end{verbatim}
\begin{itemize}
\item Time Complexity: $O(V^3)$
\item Auxiliary Space: $O(V^2)$
\end{itemize}

\end{document}